---
title: "analysis"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(
  fig.path = "out/",  
  fig.ext = "pdf",
  fig.height = 12,
  fig.width = 16,
  dpi = 300,                     # high resolution for non-vector formats
  echo = TRUE,                  
  warning = TRUE,               
  message = TRUE                
)
```

# Source and preprocess

```{r source & load}
#---- Run scripts ----

source("src/setup.R")
source("src/functions.R")

#---- MeND data ----

ds <- # dataset
  
  readRDS("data/who_hcw_survey_clean.rds")

cb <- # codebook

  read.csv("ext/data_mapping.csv", na.strings = "")

#---- External data ----

who_n_loc <- 
  
  readRDS("ext/who_n_loc.rds")

gps_eu_n <- # N GPs by country
  
  readxl::read_xlsx("ext/GP-JQ-EURO_May2026.xlsx") |> 
  janitor::clean_names()

pop_eu_n <- # EU population (for density)
  
  read_csv("ext/migr_pop3ctb_page_linear_2_0.csv") |> 
  janitor::clean_names() |> 
  select(geopolitical_entity_reporting, obs_value)

expend <- # health expenditure
  
  read_csv("ext/hlth_sha11_hp_page_linear_2_0.csv") |> 
  janitor::clean_names() |> 
  select(geopolitical_entity_reporting, obs_value)

```

```{r var lists}
#---- Descriptive level 1 ----

scdm <- c("scdm_1_rec",
          "scdm_2_rec",
          "work_1",
          "work_7")

mh_dic <- c("phq_co", "gad_co")
mh_sc <- c("phqads_sc", "phqads_z")
named_mh <- c(
  "Depressive symptoms" = "phq_co",
  "Anxiety symptoms" = "gad_co"
)
named_mh_sc <- c("Anxiety and depressive symptoms" = "phqads_sc",
                 "Anxiety and depressive symptoms (Z score)" = "phqads_z")

demands <- c("work_8", "work_11", "work_12", "work_23", "work_24") 
demands_num <- c(paste0(demands, "_num"))[c(1, 2, 4, 5)] # work_12 is binary
demands_dic <- c(paste0(demands, "_dic")) 
named_demands <- c(
  "Extended working hours"    = "work_8_dic",
  "Frequent night shifts"    = "work_11_dic",
  "Regular shift work" = "work_12_dic",
  "Tight deadlines" = "work_23_dic",
  "Angry patients or relatives" = "work_24_dic")

hazards <- c("work_30", "work_31", "work_32", "work_33")
hazards_dic <- c(paste0(hazards, "_dic"))
hazards_num <- c(paste0(hazards, "_num"))
named_hazards <- c(
  "Sexual harassment"    = "work_30_dic",
  "Violent threats"    = "work_31_dic",
  "Physical violence" = "work_32_dic",
  "Bullying" = "work_33_dic"
)

protocols <- c("work_25", "work_26", "work_27", "work_28", "work_29")
protocols_dic <- c(paste0(protocols, "_dic"))
named_protocols <- c(
  "Work council or union" = "work_25_dic",
  "Employee feedback"    = "work_26_dic",
  "Stress action plan"    = "work_27_dic",
  "Harassment protocol" = "work_28_dic",
  "Violence protocol" = "work_29_dic"
)

resources <- c("work_19", "work_20", "work_21", "work_22")
resources_dic <- c(paste0(resources, "_dic"))
resources_num <- c(paste0(resources, "_num"))
named_resources <- c(
  "Influence over work" = "work_19_dic",
  "Influence over work (breaks)"    = "work_20_dic",
  "Colleague support"    = "work_21_dic",
  "Support from superiors" = "work_22_dic"
)

all_wc <- c(demands_dic, hazards_dic, resources_dic) # protocols out
all_wc_bis <- c(demands, hazards, resources) # protocols out

#---- Descriptive level 2 ----

macro_vv <- c("gp_100k", "gp_prop", "euro_per_inh")

#---- For  MAIN analyses ----

wc_comp <- c("wc_demands", "wc_hazards", "wc_resources")

wc_comp_z <- c("wc_demands_z", "wc_hazards_z", "wc_resources_z")

confs_mh <- 
  
  c("scdm_1_rec_unord", "scdm_2_rec", "wc_demands", "wc_hazards", "wc_resources")

confs_wc  <- c("scdm_1_rec", "scdm_2_rec")
```

```{r process macro}
#---- Clean and harmonise GP counts ----
gps_n <-
  gps_eu_n |>
  mutate(country = if_else(country == "Czech Republic", "Czechia", country)) |>
  filter(country %in% ds$loc_2) |>
  rename(loc_2 = country, gp_n = general_practitioners) |>
  mutate(year_gp = as.integer(years), gp_n = as.numeric(gp_n)) |>
  filter(!is.na(loc_2), !is.na(year_gp), !is.na(gp_n)) |>
  group_by(loc_2) |>
  slice_max(year_gp, n = 1, with_ties = FALSE) |>
  ungroup() |>
  select(loc_2, gp_n)

pop_eu_n <-
  pop_eu_n |>
  rename(loc_2 = geopolitical_entity_reporting, pop_n = obs_value) |>
  filter(loc_2 %in% ds$loc_2)

expend <-
  expend |>
  rename(loc_2 = geopolitical_entity_reporting, euro_per_inh = obs_value)

#---- Build macro dataset ----
macro_ds <-
  gps_n |>
  left_join(pop_eu_n, by = "loc_2") |>
  mutate(gp_100k = gp_n / pop_n * 100000) |>
  left_join(
    who_n_loc |>
      filter(work_2 == "Doctor") |>
      select(loc_2, value) |>
      rename(doc_n = value),
    by = "loc_2"
  ) |>
  mutate(gp_prop = gp_n / doc_n * 100) |>
  left_join(expend, by = "loc_2")

#---- Obtain coverage by country ----

country_coverage <- 
  
  ds |> 
  filter(specialty_cat == "Primary Care") |> 
  summarise(n = n(), .by = "loc_2") |> 
  left_join(macro_ds, by = "loc_2") |> 
  mutate(coverage = n/gp_n*100) |> 
  select(loc_2, n, gp_n, coverage)
```

```{r process ds}
#---- 1. Shorten long factor labels ----
ds <- ds |>
  mutate(
    work_7 = fct_recode(
      work_7,
      "Public and private sectors" =
        "I work in both the public sector and in the private sector (or I work in the public sector and I am self-employed)",
      "Private sector only" =
        "I work in the private healthcare sector only (I do not work in the public sector)",
      "Public sector only" =
        "I work in the public healthcare sector only (I do not work in the private sector, nor am I self-employed)"
    ),
    scdm_1 = fct_recode(scdm_1, "Under 20" = "Less than 20 years old"),
    work_5 = fct_recode(
      work_5,
      "Extra-hospital emergencies" = "Emergency care outside hospital (e.g., ambulance)",
      "In-hospital care"           = "Inpatient care (hospital services)",
      "Outpatient primary care"    = "Primary care (outpatient services)",
      "Outpatient secondary care"  = "Secondary care (outpatient services)"
    ),
    work_8 = fct_recode(
      work_8,
      "< 11 h"   = "Less than 11 hours",
      "11-20 h"  = "11-20 hours",
      "21-30 h"  = "21-30 hours",
      "31-40 h"  = "31-40 hours",
      "41-50 h"  = "41-50 hours",
      "51-60 h"  = "51-60 hours",
      "61-70 h"  = "61-70 hours",
      "> 70 h"   = "More than 70 hours"
    ),
    work_11 = fct_recode(
      work_11,
      "Daily"             = "Every day",
      "Almost daily"      = "At least 4 days a week, but not every day",
      "2-3 days/week"     = "2 or 3 days a week",
      "Almost weekly"     = "Almost every week",
      "1-2 times/month"   = "Once or twice a month",
      "Less than monthly" = "Less than once a month",
      "Never/Rarely"      = "Never or very rarely (e.g., once or twice a year)"
    )
  )

#---- 2. Derived sociodemographic variables ----
ds <- ds |>
  mutate(
    work_5_rec = fct_collapse(
      work_5,
      "Inpatient"  = c("Emergency department", "In-hospital care"),
      "Outpatient" = c("Outpatient primary care", "Outpatient secondary care"),
      "Other"      = c("Extra-hospital emergencies", "Other")
    ),
    scdm_5_dic = factor(
      case_when(
        str_detect(scdm_5, "with children")    ~ "Has children",
        str_detect(scdm_5, "without children") ~ "No children"
      ),
      levels = c("Has children", "No children")
    ),
    work_3_rec = fct_collapse(
      work_3,
      "5 years or less"    = c("Less than 1 year", "1-3 years", "4-5 years"),
      "More than 10 years" = c("11-20 years", "21-30 years", "31-40 years",
                                "More than 40 years")
    ),
    work_4_dic = factor(
      if_else(work_4 == "One", "One", "More than one"),
      levels = c("One", "More than one")
    ),
    work_16_dic = factor(
      fct_recode(work_16, "Yes" = "Temporary contract", "No" = "Fixed contract"),
      levels = c("No", "Yes")
    ),
    # Age: unordered factor with 50-59 as reference (lowest turnover)
    scdm_1_rec_unord = factor(
      scdm_1_rec,
      levels  = c("50-59", "<30", "30-39", "40-49", "60+"),
      ordered = FALSE
    )
  )

#---- 3. Working condition dichotomies ----
ds <- ds |>
  mutate(
    # Demands: hours and nights
    work_8_dic  = factor(
      if_else(as.numeric(work_8) > 5, "Yes", "No"),    # > 50 h/week
      levels = c("No", "Yes")
    ),
    work_11_dic = factor(
      if_else(as.numeric(work_11) <= 5, "Yes", "No"),  # >= almost weekly
      levels = c("No", "Yes")
    ),
    work_11     = fct_rev(work_11),
    work_12_dic = factor(work_12, levels = c("No", "Yes")),
    # Demands: psychosocial (work_23, work_24)
    across(
      all_of(c("work_23", "work_24")),
      ~ factor(
        fct_collapse(
          .x,
          "Yes" = c("Around 1/4 of the time", "Around half of the time",
                    "Around 3/4 of the time", "Almost all of the time",
                    "All of the time"),
          "No"  = c("Never", "Almost never")
        ),
        levels = c("No", "Yes")
      ),
      .names = "{.col}_dic"
    ),
    # Hazards (violence)
    across(
      all_of(hazards),
      ~ factor(
        fct_collapse(
          .x,
          "Yes" = c("Yes, daily", "Yes, weekly", "Yes, monthly", "Yes, a few times"),
          "No"  = "No"
        ),
        levels = c("No", "Yes")
      ),
      .names = "{.col}_dic"
    ),
    # Protocols: drop "Do not know", relevel to No
    across(
      all_of(protocols),
      ~ fct_relevel(fct_recode(.x, NULL = "Do not know"), "No"),
      .names = "{.col}_dic"
    ),
    # Resources
    across(
      all_of(resources),
      ~ factor(
        fct_collapse(
          .x,
          "Yes" = c("Always", "Often"),
          "No"  = c("Sometimes", "Seldom", "Never / Hardly ever")
        ),
        levels = c("No", "Yes")
      ),
      .names = "{.col}_dic"
    ),
    # Work-life balance
    work_34_dic = factor(
      if_else(as.numeric(work_34) > 3, "Yes", "No"),
      levels = c("No", "Yes")
    )
  )

#---- 4. Numeric versions of WC items (for composites) ----
ds <- ds |>
  mutate(
    across(
      all_of(hazards),
      ~ case_when(
        .x == "No"               ~ 0,
        .x == "Yes, a few times" ~ 1,
        .x == "Yes, monthly"     ~ 2,
        .x == "Yes, weekly"      ~ 3,
        .x == "Yes, daily"       ~ 4,
        TRUE                     ~ NA_real_
      ),
      .names = "{.col}_num"
    ),
    across(
      all_of(demands) & !all_of("work_12"),  # work_12 is already binary
      ~ as.numeric(.x) - 1,
      .names = "{.col}_num"
    ),
    across(
      all_of(resources),
      ~ as.numeric(.x) - 1,
      .names = "{.col}_num"
    )
  )

#---- 5. WC composites and z-scores ----
ds <- ds |>
  mutate(
    wc_demands   = rowSums(across(all_of(demands_num)),  na.rm = FALSE),
    wc_hazards   = rowSums(across(all_of(hazards_num)),  na.rm = FALSE),
    wc_resources = rowSums(                              # reverse-scored
      across(all_of(resources_num), ~ max(.x, na.rm = TRUE) - .x),
      na.rm = FALSE
    )
  ) |>
  mutate(
    across(
      all_of(wc_comp),
      ~ as.numeric(scale(.x)),
      .names = "{.col}_z"
    )
  )

#---- 6. Mental health exposures ----
ds <- ds |>
  mutate(
    phqads_sc = phq_sc + gad_sc,
    phqads_z  = as.numeric(scale(phqads_sc))
  )

#---- 7. Primary outcome ----
ds <- ds |>
  mutate(
    work_18_poi = as.integer(if_else(work_18 == "Yes", 1L, 0L))
  )

#---- 8. Join macro variables ----
ds <- left_join(ds, macro_ds |> select(loc_2, all_of(macro_vv)), by = "loc_2")
```

```{r grids}
#---- Main model grids ----
grid_gp_adj <- bind_rows(
  tibble(exposure_var = c("phqads_z", "phq_co", "gad_co"),
         confounder_key = "mh"),
  tibble(exposure_var = c("wc_demands_z", "wc_hazards_z", "wc_resources_z"),
         confounder_key = "wc")
) |>
  mutate(
    outcome_var = "work_18_poi",
    profession  = "GP",
    cluster_var = "loc_2",
    family_type = "poisson",
    model_label = "Fully adjusted"
  )

grid_gp <- bind_rows(
  grid_gp_adj,
  grid_gp_adj |> mutate(model_label = "Crude", confounder_key = NA)
)

grid_nongp <- grid_gp |> mutate(profession = "Non-GP")

#---- Mediation grids ----
grid_med_gp <- expand_grid(
  exposure_var = c("wc_demands_z", "wc_hazards_z", "wc_resources_z"),
  mediator_var = c("phqads_z", "phq_co", "gad_co")
) |>
  mutate(
    outcome_var    = "work_18_poi",
    profession     = "GP",
    cluster_var    = "loc_2",
    confounder_key = "wc"
  )

grid_med_nongp <- grid_med_gp |> mutate(profession = "Non-GP")
```

```{r filters}
ds_gp <- ds |> filter(specialty_cat == "Primary Care")
ds_nongp <- ds |> filter(specialty_cat != "Primary Care")
```

```{r preprocess OLD, eval = FALSE}

# Some responses are too long

ds <- 
  
  ds  |> 
  mutate(work_7 = fct_recode(
    work_7,
    "Public and private sectors" = 
      "I work in both the public sector and in the private sector (or I work in the public sector and I am self-employed)",
    "Private sector only" = 
      "I work in the private healthcare sector only (I do not work in the public sector)",
    "Public sector only" = 
      "I work in the public healthcare sector only (I do not work in the private sector, nor am I self-employed)"
  ))


ds <- 
  
  ds |> 
  mutate(scdm_1 = fct_recode(scdm_1, "Under 20" = "Less than 20 years old"))


ds <- 
  
  ds |> 
  mutate(work_5 = fct_recode(
    work_5,
    "Extra-hospital emergencies" = 
      "Emergency care outside hospital (e.g., ambulance)",
    "In-hospital care" = 
      "Inpatient care (hospital services)",
    "Outpatient primary care" = 
      "Primary care (outpatient services)",
    "Outpatient secondary care" =
      "Secondary care (outpatient services)"
  ))


ds <- 
  
  ds |> 
  mutate(work_5_rec = fct_collapse( 
    work_5,
    "Inpatient" = c("Emergency department",
                    "In-hospital care"),
    "Outpatient" = c("Outpatient primary care",
                     "Outpatient secondary care"),
    "Other" = c("Extra-hospital emergencies",
                "Other")
    )
  )

ds <- 
  
  ds |> 
  mutate(work_11 = fct_recode(
    work_11,
    "Daily" = 
      "Every day",
    "Almost daily" = 
      "At least 4 days a week, but not every day",
    "2-3 days/week" = 
      "2 or 3 days a week",
    "Almost weekly" = 
      "Almost every week",
    "1-2 times/month" =
      "Once or twice a month",
    "Less than monthly" =
    "Less than once a month",
    "Never/Rarely" = 
      "Never or very rarely (e.g., once or twice a year)"
  ))


ds <- 
  
  ds |> 
  mutate(work_8 = fct_recode(
    work_8,
    "< 11 h" = 
      "Less than 11 hours",
    "11-20 h" = 
      "11-20 hours",
    "21-30 h" = 
      "21-30 hours",
    "31-40 h" = 
      "31-40 hours",
    "41-50 h" =
      "41-50 hours",
    "51-60 h" = 
      "51-60 hours",
    "61-70 h" = 
      "61-70 hours",
    "> 70 h" = 
      "More than 70 hours"
  ))

# Some variables and exposures require processing

ds <- 
  
  ds |> 
  
  # Recode household composition 
  
  mutate(scdm_5_dic = case_when(
    str_detect(scdm_5, "with children") ~ "Has children",
    str_detect(scdm_5, "without children") ~ "No children",
    TRUE ~ NA
    ),
    scdm_5_dic = factor(scdm_5_dic,
                        levels = c("Has children", "No children"),
                        ordered = FALSE)
  ) |>
  
  # recode length of service
  
  mutate(work_3_rec = fct_collapse(
    work_3,
    "5 years or less" = c("Less than 1 year", "1-3 years", "4-5 years"),
    "More than 10 years" = c("11-20 years", "21-30 years", "31-40 years",
                             "More than 40 years")
    )
  ) |> 
  
  # Recode number of job positions
  
  mutate(
    work_4_dic = if_else(work_4 == "One",
                         "One",
                         "More than one"),
    work_4_dic = factor(work_4_dic,
                        levels = c("One", "More than one"))
  ) |> 
  
  # Recode weekly working hours
  
  mutate(
    work_8_dic = case_when(as.numeric(work_8) <= 5 ~ "No", # <= 50h
                           as.numeric(work_8) > 5 ~ "Yes"), # > 50h
    work_8_dic = factor(work_8_dic, levels = c("No", "Yes"))
  ) |> 
  
  # Recode frequency of night work
  
  mutate(
    work_11_dic = 
      case_when(as.numeric(work_11) > 5 ~ "No", # 1 or 2x a month or less
                as.numeric(work_11) <= 5 ~ "Yes"), # Almost every week and more
    work_11_dic = factor(work_11_dic, levels = c("No", "Yes")),
    work_11 = fct_rev(work_11)
  ) |> 
  
  # name var for shifts
  
  mutate(work_12_dic = factor(work_12, levels = c("No", "Yes"))
  ) |> 
  
  # create dichotomous var for contract type
  
  mutate(work_16_dic = fct_recode(work_16,
                                  "Yes" = "Temporary contract",
                                  "No" = "Fixed contract"),
         work_16_dic = factor(work_16_dic,
                              levels = c("No", "Yes"),
                              ordered = FALSE)
  ) |> 
  
  # recode managing patient anger
  
  mutate(
    across(
      all_of(c("work_23", "work_24")),
      ~ factor(
        fct_collapse(.x,
                     "Yes" = c("Around 1/4 of the time",
                               "Around half of the time",
                               "Around 3/4 of the time",
                               "Almost all of the time",
                               "All of the time"),
                     "No" = c("Never",
                              "Almost never")),
                     levels = c("No", "Yes"),
                     ordered = FALSE),
        .names = "{.col}_dic"
        )
      ) |> 
  
  # recode violence
  
   mutate(
    across(
      all_of(hazards),
      ~ factor(
        fct_collapse(.x,
                     "Yes" = c("Yes, daily", 
                             "Yes, weekly", 
                             "Yes, monthly", 
                             "Yes, a few times"),
                     "No" = "No"
        ),
        levels = c("No", "Yes"), ordered = FALSE
      ),
      .names = "{.col}_dic"
    )
  ) |>
  
  # Recode protocols, drop "Do not know", ensure factor levels
 
   mutate(
    across(
      all_of(protocols),
      ~ fct_recode(.x, NULL = "Do not know"),
      .names = "{.col}_dic"
    )
  ) |> 

  # relevel protocols

  mutate(
    across(
      all_of(protocols_dic),
      ~ fct_relevel(.x, "No")
    )
  ) |> 
  
  # recode job resources
  
  mutate(
    across(
      all_of(c("work_19", "work_20", "work_21", "work_22")),
      ~ factor(
        fct_collapse(.x,
                     "No" = c("Sometimes", "Seldom", "Never / Hardly ever"),
                     "Yes" = c("Always", "Often"),
        ),
        levels = c("No", "Yes"), ordered = FALSE
      ),
      .names = "{col}_dic"
    )
  ) |> 
  
  # recode work-life balance
  
  mutate(
    work_34_dic = case_when(as.numeric(work_34) <= 3 ~ "No",
                            as.numeric(work_34) > 3 ~ "Yes"),
    work_34_dic = factor(work_34_dic, levels = c("No", "Yes"))
  ) |> 
  
  # create numeric variables
  
  mutate(
    across(
      all_of(hazards),
      ~ case_when(
        .x == "No"                ~ 0,
        .x == "Yes, a few times"  ~ 1,
        .x == "Yes, monthly"      ~ 2,
        .x == "Yes, weekly"       ~ 3,
        .x == "Yes, daily"        ~ 4,
        TRUE                      ~ NA_real_
        ),
      .names = "{.col}_num"
      )
    ) |> 
  mutate(
    across(
      all_of(demands) & !all_of("work_12"), # work_12 is binary
      \(x) as.numeric(x) - 1, # 0 as reference
    .names = "{.col}_num")) |> 
  mutate(
    across(
      all_of(resources),
      \(x) as.numeric(x) - 1, # 0 as reference
    .names = "{.col}_num")) |> 
  
  # create poisson outcome variables

  mutate(
    work_18_poi = if_else(work_18 == "Yes", 1, 0) |> as.integer()
  ) |> 
  
  mutate(
    phqads_sc = phq_sc + gad_sc,
    phqads_z = 
      (phqads_sc - mean(phqads_sc, na.rm = TRUE)) / sd(phqads_sc, na.rm = TRUE)
  )

# Composite WC score

ds <- 
  
  ds |> 
  mutate(
    wc_demands  = rowSums(across(all_of(demands_num)),  na.rm = FALSE),
    wc_hazards  = rowSums(across(all_of(hazards_num)),  na.rm = FALSE),
    wc_resources = rowSums(
      across(all_of(resources_num), ~ max(.x, na.rm = TRUE) - .x), # reverse
      na.rm = FALSE
    )
  ) |> 
  mutate(
    wc_demands_z  = 
      (wc_demands  - mean(wc_demands,  na.rm = TRUE)) / 
      sd(wc_demands,  na.rm = TRUE),
    wc_hazards_z  = 
      (wc_hazards  - mean(wc_hazards,  na.rm = TRUE)) / 
      sd(wc_hazards,  na.rm = TRUE),
    wc_resources_z = 
      (wc_resources - mean(wc_resources, na.rm = TRUE)) / 
      sd(wc_resources, na.rm = TRUE)
  )

ds <- 
  
  ds |> 
  mutate(
    scdm_1_rec_num = case_when(scdm_1_rec == "<30"   ~ "1",
                               scdm_1_rec == "30-39" ~ "2",
                               scdm_1_rec == "40-49" ~ "3",
                               scdm_1_rec == "50-59" ~ "4",
                               scdm_1_rec == "60+"   ~ "5",
                               TRUE                  ~ scdm_1_rec
                               ) |> as.integer(),
    scdm_1_rec_unord = factor(scdm_1_rec, 
                              ordered = FALSE,
                              levels = c("50-59", # reference, lower ITL
                                         "<30",
                                         "30-39",
                                         "40-49",
                                         "60+"))
    )
```

# Coverage

```{r coverage}
country_coverage |> 
  ggplot(aes(x = loc_2, y = coverage)) +
  geom_col() +
  geom_text(aes(label = n), vjust = -0.5)
```

# Macro

```{r gp 100k}
plot_df <- ds |>
  filter(specialty_cat == "Primary Care") |>
  left_join(macro_ds, by = "loc_2") |>
  drop_na(gp_100k) |> 
  mutate(loc_2 = fct_reorder(loc_2, gp_100k))

# linear rescale of gp_100k 
rng_phq <- range(plot_df$phq_sc, na.rm = TRUE)
rng_gp  <- range(plot_df$gp_100k, na.rm = TRUE)
b <- diff(rng_phq) / diff(rng_gp)
a <- rng_phq[1] - rng_gp[1] * b

p1 <- 
  
  plot_df |>
  filter(!loc_2 %in% c("Poland", "Portugal")) |> 
  group_by(loc_2, gp_100k) |>
  summarise(mean_phq = mean(phq_sc, na.rm = TRUE), n = n(), .groups = "drop") |>
  ggplot(aes(x = gp_100k, y = mean_phq)) +
  geom_point(aes(size = n)) +
  geom_smooth(method = "lm", se = TRUE) +
  ggrepel::geom_text_repel(aes(label = loc_2), size = 3) +
  theme_minimal()

# linear rescale of gp_100k
rng_gad <- range(plot_df$gad_sc, na.rm = TRUE)
rng_gp  <- range(plot_df$gp_100k, na.rm = TRUE)
b <- diff(rng_gad) / diff(rng_gp)
a <- rng_gad[1] - rng_gp[1] * b

p2 <- 
  
  plot_df |>
  filter(!loc_2 %in% c("Poland", "Portugal")) |> 
  group_by(loc_2, gp_100k) |>
  summarise(mean_gad = mean(gad_sc, na.rm = TRUE), n = n(), .groups = "drop") |>
  ggplot(aes(x = gp_100k, y = mean_gad)) +
  geom_point(aes(size = n)) +
  geom_smooth(method = "lm", se = TRUE) +
  ggrepel::geom_text_repel(aes(label = loc_2), size = 3) +
  theme_minimal()

p1/p2
```

```{r gp prop}
plot_df <- ds |>
  filter(specialty_cat == "Primary Care") |>
  left_join(macro_ds, by = "loc_2") |>
  drop_na(gp_prop) |> 
  mutate(loc_2 = fct_reorder(loc_2, gp_prop))

# linear rescale of gp_prop 
rng_phq <- range(plot_df$phq_sc, na.rm = TRUE)
rng_gp  <- range(plot_df$gp_prop, na.rm = TRUE)
b <- diff(rng_phq) / diff(rng_gp)
a <- rng_phq[1] - rng_gp[1] * b

p1 <- 
  
  plot_df |>
  filter(!loc_2 %in% c("Poland", "Portugal")) |> 
  group_by(loc_2, gp_prop) |>
  summarise(mean_phq = mean(phq_sc, na.rm = TRUE), n = n(), .groups = "drop") |>
  ggplot(aes(x = gp_prop, y = mean_phq)) +
  geom_point(aes(size = n)) +
  geom_smooth(method = "lm", se = TRUE) +
  ggrepel::geom_text_repel(aes(label = loc_2), size = 3) +
  theme_minimal()

# linear rescale of gp_prop
rng_gad <- range(plot_df$gad_sc, na.rm = TRUE)
rng_gp  <- range(plot_df$gp_prop, na.rm = TRUE)
b <- diff(rng_gad) / diff(rng_gp)
a <- rng_gad[1] - rng_gp[1] * b

p2 <- 
  
  plot_df |>
  filter(!loc_2 %in% c("Poland", "Portugal")) |> 
  group_by(loc_2, gp_prop) |>
  summarise(mean_gad = mean(gad_sc, na.rm = TRUE), n = n(), .groups = "drop") |>
  ggplot(aes(x = gp_prop, y = mean_gad)) +
  geom_point(aes(size = n)) +
  geom_smooth(method = "lm", se = TRUE) +
  ggrepel::geom_text_repel(aes(label = loc_2), size = 3) +
  theme_minimal()

p1/p2
```

```{r expend}
plot_df <- ds |>
  filter(specialty_cat == "Primary Care") |>
  left_join(macro_ds, by = "loc_2") |>
  drop_na(euro_per_inh) |> 
  mutate(loc_2 = fct_reorder(loc_2, euro_per_inh))

# linear rescale of euro_per_inh 
rng_phq <- range(plot_df$phq_sc, na.rm = TRUE)
rng_gp  <- range(plot_df$euro_per_inh, na.rm = TRUE)
b <- diff(rng_phq) / diff(rng_gp)
a <- rng_phq[1] - rng_gp[1] * b

p1 <- 
  
  plot_df |>
  group_by(loc_2, euro_per_inh) |>
  summarise(mean_phq = mean(phq_sc, na.rm = TRUE), n = n(), .groups = "drop") |>
  ggplot(aes(x = euro_per_inh, y = mean_phq)) +
  geom_point(aes(size = n)) +
  geom_smooth(method = "lm", se = TRUE) +
  ggrepel::geom_text_repel(aes(label = loc_2), size = 3) +
  theme_minimal()

# linear rescale of euro_per_inh
rng_gad <- range(plot_df$gad_sc, na.rm = TRUE)
rng_gp  <- range(plot_df$euro_per_inh, na.rm = TRUE)
b <- diff(rng_gad) / diff(rng_gp)
a <- rng_gad[1] - rng_gp[1] * b

p2 <- 
  
  plot_df |>
  group_by(loc_2, euro_per_inh) |>
  summarise(mean_gad = mean(gad_sc, na.rm = TRUE), n = n(), .groups = "drop") |>
  ggplot(aes(x = euro_per_inh, y = mean_gad)) +
  geom_point(aes(size = n)) +
  geom_smooth(method = "lm", se = TRUE) +
  ggrepel::geom_text_repel(aes(label = loc_2), size = 3) +
  theme_minimal()

p1/p2
```

## turnover

```{r}
plot_df <- ds |>
  filter(specialty_cat == "Primary Care") |>
  left_join(macro_ds, by = "loc_2") |>
  drop_na(euro_per_inh) |> 
  mutate(loc_2 = fct_reorder(loc_2, euro_per_inh))


p1 <- 
  
  plot_df |>
  group_by(loc_2, euro_per_inh) |>
  summarise(
    work_18_prop = mean(work_18 == "Yes", na.rm = TRUE),
    n = n(), 
    .groups = "drop") |>
  ggplot(aes(x = euro_per_inh, y = work_18_prop)) +
  geom_point(aes(size = n)) +
  geom_smooth(method = "lm", se = TRUE) +
  ggrepel::geom_text_repel(aes(label = loc_2), size = 3) +
  theme_minimal()

p2 <- 
  
  plot_df |>
  group_by(loc_2, gp_100k) |>
  summarise(
    work_18_prop = mean(work_18 == "Yes", na.rm = TRUE),
    n = n(), 
    .groups = "drop") |>
  ggplot(aes(x = gp_100k, y = work_18_prop)) +
  geom_point(aes(size = n)) +
  geom_smooth(method = "lm", se = TRUE) +
  ggrepel::geom_text_repel(aes(label = loc_2), size = 3) +
  theme_minimal()

p3 <- 
  
  plot_df |>
  group_by(loc_2, gp_prop) |>
  summarise(
    work_18_prop = mean(work_18 == "Yes", na.rm = TRUE),
    n = n(), 
    .groups = "drop") |>
  ggplot(aes(x = gp_prop, y = work_18_prop)) +
  geom_point(aes(size = n)) +
  geom_smooth(method = "lm", se = TRUE) +
  ggrepel::geom_text_repel(aes(label = loc_2), size = 3) +
  theme_minimal()



p1/p2/p3

```

# Models

## Main

```{r run main}
run_grid <- function(grid, data) {
  pmap(grid, function(outcome_var, exposure_var, confounder_key,
                      profession, cluster_var, family_type, model_label) {
    confounders <- case_when(
      model_label == "Crude"   ~ list(NULL),
      confounder_key == "mh"   ~ list(confs_mh),
      confounder_key == "wc"   ~ list(confs_wc)
    )[[1]]
    run_adj_glm_v2(
      data            = data,
      outcome_var     = outcome_var,
      exposure_var    = exposure_var,
      confounder_list = confounders,
      profession      = profession,
      cluster_var     = cluster_var,
      family_type     = family_type,
      model_label     = model_label
    )
  })
}

results_gp    <- run_grid(grid_gp, ds_gp)
results_nongp <- run_grid(grid_nongp, ds_nongp)

results_main <-
  list(
    map_dfr(results_gp,    "multilevel"),
    map_dfr(results_gp,    "robust"),
    map_dfr(results_nongp, "multilevel"),
    map_dfr(results_nongp, "robust")
  ) |>
  bind_rows()

saveRDS(results_main, "out/models/summary/results_main.rds")
```

# Mediation

```{r run med}
#---- Run mediation ----
run_grid_med <- function(grid, data, sims = 1000) {
  pmap(grid, function(exposure_var, mediator_var, outcome_var,
                      profession, cluster_var, confounder_key) {
    confounders <- if (confounder_key == "mh") confs_mh else confs_wc
    run_mediation(
      data            = data,
      exposure_var    = exposure_var,
      mediator_var    = mediator_var,
      outcome_var     = outcome_var,
      confounder_list = confounders,
      profession      = profession,
      cluster_var     = cluster_var,
      sims            = sims
    )
  })
}

results_med_gp    <- run_grid_med(grid_med_gp,    ds_gp)
results_med_nongp <- run_grid_med(grid_med_nongp, ds_nongp)

results_med <-
  list(
    results_med_gp,
    results_med_nongp
  ) |>
  map(~ bind_rows(.x)) |>
  bind_rows()

saveRDS(results_med, "out/models/summary/results_med.rds")
```

```{r med tbl}
#---- GP mediation table ----
fmt_ci <- function(est, lo, hi, mult = 100, digits = 1) {
  sprintf(paste0("%.", digits, "f (%.", digits, "f, %.", digits, "f)"),
          est * mult, lo * mult, hi * mult)
}

med_tbl_gp_df <-
  results_med |>
  filter(Profession == "GP") |>
  mutate(
    Exposure = factor(
      Exposure,
      levels = c("wc_demands_z", "wc_hazards_z", "wc_resources_z"),
      labels = c("Job demands", "Job hazards", "Workplace resources")
    ),
    Mediator = factor(
      Mediator,
      levels = c("phqads_z", "phq_co", "gad_co"),
      labels = c(
        "Anxiety & depressive symptoms (PHQ-ADS)",
        "Probable depression (PHQ-9)",
        "Probable anxiety (GAD-7)"
      )
    )
  ) |>
  arrange(Exposure, Mediator) |>
  transmute(
    Exposure,
    Mediator,
    N        = format(N, big.mark = ","),
    Indirect = fmt_ci(ACME,  ACME_CI_l,  ACME_CI_u),
    Direct   = fmt_ci(ADE,   ADE_CI_l,   ADE_CI_u),
    Total    = fmt_ci(Total, Total_CI_l, Total_CI_u),
    Pct_med  = fmt_ci(Prop_mediated, Prop_med_CI_l, Prop_med_CI_u)
  )

med_tbl_gp <-
  med_tbl_gp_df |>
  flextable() |>
  set_header_labels(
    Exposure = "Working condition",
    Mediator = "Mediator (mental health)",
    N        = "N",
    Indirect = "Indirect (ACME)",
    Direct   = "Direct (ADE)",
    Total    = "Total",
    Pct_med  = "% mediated"
  ) |>
  add_header_row(
    values    = c("", "", "", "Risk difference (percentage points)", ""),
    colwidths = c(1, 1, 1, 3, 1),
    top       = TRUE
  ) |>
  merge_v(j = "Exposure") |>
  valign(j = "Exposure", valign = "top", part = "body") |>
  align(i = 1, align = "center", part = "header") |>
  align(j = c("N", "Indirect", "Direct", "Total", "Pct_med"),
        align = "center", part = "all") |>
  theme_booktabs() |>
  fontsize(size = 9, part = "all") |>
  autofit() |>
  add_footer_lines(c(
    "Effects are absolute differences in the predicted probability of intention to leave the profession (work_18), in percentage points, per 1-SD increase in the working-condition composite. All composites oriented so that higher scores indicate more adverse conditions (workplace resources reverse-scored). Models adjusted for age and sex with a country random intercept; mediation estimated by quasi-Bayesian approximation (1000 simulations).",
    "ACME, average causal mediation effect (indirect path through mental health); ADE, average direct effect. PHQ-ADS, summed PHQ-9 and GAD-7 symptom score (per SD); PHQ-9, depression caseness (cutoff ≥10); GAD-7, anxiety caseness (cutoff ≥10)."
  ))

save_as_docx(med_tbl_gp, path = "out/tables/med_tbl_gp.docx")
```

Important distinction here: the closest value isn't an exact match to
anything in results_main, and it's worth being precise about why. The
scales differ. Your main model row 4 gives a prevalence ratio of 1.49 —
multiplicative, "demands raise turnover risk by 49% per SD." The
mediation Total of \~4.0 pp is additive, "demands raise the absolute
probability of turnover by 4 percentage points per SD." These describe
the same association but on different scales, so they will never be
numerically equal. Which mediation row corresponds to that PR. The
relevant comparison is the mediation row with the same exposure and the
same confounder set. Your main model row 4 (wc_demands_z, GP, fully
adjusted with confs_wc = age + sex) corresponds to the demands rows in
the mediation table. But there are three of them (4.0, 4.7, 4.6), one
per mediator — and here's the subtle point: the Total effect should be
identical across all three, because the total effect of demands on
turnover does not depend on which mediator you chose. The small
differences (4.0 vs 4.7 vs 4.6) come entirely from the analytic sample
changing via listwise deletion — each mediation model drops rows missing
that particular mediator, and PHQ-9/GAD-7 caseness have different
missingness than the continuous PHQ-ADS. So the honest answer: all three
are estimating the same total effect of demands; they differ only
because of differing N. The PHQ-ADS row (4.0 pp, N = 8027) is on the
largest, most complete sample, so it's the most stable estimate of the
demands total effect. To actually verify the two analyses agree, don't
compare the PR to the pp directly. Instead, convert. Your main model
also reports a baseline risk you can use: for a binary exposure like
phq_co you have Risk_Unexp = 6.4% and Risk_Exp = 14.4%, a difference of
8 pp on the response scale. For the continuous wc_demands_z the PR is
1.49, and at a baseline turnover risk of roughly 8–9% in GPs, a 49%
multiplicative increase per SD works out to roughly 4 pp on the absolute
scale — which lands right on your mediation Total of 4.0. That
back-of-envelope reconciliation is the check that the two pipelines are
telling a consistent story. One caveat worth stating in the methods: the
mediation Total comes from a glmer Poisson outcome model with the
mediator also in the model structure, whereas the main-model PR comes
from a Poisson model without the mediator. They won't reconcile to
arbitrary precision, but they should be in the same ballpark, and they
are.
